Natural Selection Across Three Generations of Americans

### 2024-02-23 hrs-selection.qmd

@@

$$
\hat{\beta}_{TRUE} = \hat{\beta}_{PGS}\sqrt{\frac{h^2}{R^2_{PGS}}
}$$

>>>>

```{r}
#| label: top-PGS
#| eval: false

# Is it ok to residualise PGS on PCS vs control for them?
# Frisch-Waugh-Lovell theorem: it's the same thing... But we need PGS standardized to sd of 1
# if the regression beta is to be the covariance for the Price Equation
# When you residualise on PCs *and then restandardize* then the beta *is different* than just controlling for PCs. The difference is marginal (~1% lower in the below example). Furthermore, controlling for PCs seems to have been the standard practice in Beauchamp and in Kong et al.

# I'm not sure whether we can residualise on PCs then use the error correction because PGS true, will correlate differently with the PCs. Then again I'm not sure that the correction would be right for twin H2 since the PCs will surely only correlate with the common SNPs? I've consulted the Becker et al 2021 supplementary and it doesn't give advice on the use of PCs. I'd have to think very carefully about how to do this.

# Either we justify our approach as reasonable. Or we residualise on the PCs and go with it.
# I'm in favour of the latter, in part because explaining why the effect sizes switch sign in the error correction is awkward.

reg_simple <- lm(rlrs ~ scale(EA) ,rand$variables)
reg_pc <- lm(rlrs ~ scale(EA) + PC1_5A +PC1_5B +PC1_5C +PC1_5D +PC1_5E + 
             PC6_10A +PC6_10B +PC6_10C +PC6_10D +PC6_10E ,rand$variables)
# beta = -0.027651
reg_res_EA <- lm(`scale(EA)` ~ PC1_5A +PC1_5B +PC1_5C +PC1_5D +PC1_5E + 
             PC6_10A +PC6_10B +PC6_10C +PC6_10D +PC6_10E ,reg_pc$model)

reg_res_rlrs <- lm(rlrs ~ PC1_5A +PC1_5B +PC1_5C +PC1_5D +PC1_5E + 
             PC6_10A +PC6_10B +PC6_10C +PC6_10D +PC6_10E ,reg_pc$model)

reg_res_noscale <- lm(reg_res_rlrs$residuals ~ reg_res_EA$residuals, 
                               reg_pc$model)
# beta = -0.027651
reg2_res <- lm(reg_res_rlrs$residuals ~ scale(reg_res_EA$residuals),
                       reg_pc$model)
# beta = -0.027358

reg2_res_pgs_only <- lm(rlrs ~ scale(reg_res_EA$residuals),
                       reg_pc$model)
# beta = -0.027358

# Below I get the estimate of the interactions and their p values. 
# EA is only significant for income at p = 0.0106
# There are always traits significant after bonferroni. But usually it's autism
# Given DHJ's improvement to manuscript, I don't think we have anything more convincing to give by focusing on the top PGS.

  # the colon interaction estimates the effect separately within each group
  form_pgs_x_ed <- paste0(pgs, "*edu") |>
    map(\(x) reformulate(c(x, "edu", pcs), response = "rlrs"))
  mod_pgs_x_ed[[eth]]  <-
    map(form_pgs_x_ed, 
               \(x) svyglm_quiet(x, r_eth))

  tidy_pgs_x_ed[[eth]] <-
    map(mod_pgs_x_ed[[eth]], broom::tidy, conf.int = TRUE)
  names(tidy_pgs_x_ed[[eth]]) <- pgs
  tidy_pgs_x_ed[[eth]] <-
    list_rbind(tidy_pgs_x_ed[[eth]], names_to = "pgs") |>
    filter(grepl(':', term)) |>
    mutate(edu = str_remove(term, ".*:"))

  form_pgs_x_inc <- paste0(pgs, "*income_med") |>
    map(\(x) reformulate(c(x, "income_med", pcs), response = "rlrs"))
  mod_pgs_x_inc[[eth]]  <-
    map(form_pgs_x_inc, 
               \(x) svyglm_quiet(x, r_eth))

  tidy_pgs_x_inc[[eth]] <-
    map(mod_pgs_x_inc[[eth]], broom::tidy, conf.int = TRUE)
  names(tidy_pgs_x_inc[[eth]]) <- pgs
  tidy_pgs_x_inc[[eth]] <-
    list_rbind(tidy_pgs_x_inc[[eth]], names_to = "pgs") |>
    filter(grepl(':', term)) |>
    mutate(income_med = str_remove(term, ".*:income_med"))

  form_pgs_x_mar <- paste0(pgs, "*married") |>
    map(\(x) reformulate(c(x, "married", pcs), response = "rlrs"))
  mod_pgs_x_mar[[eth]]  <-
    map(form_pgs_x_mar, 
               \(x) svyglm_quiet(x, r_eth))

  tidy_pgs_x_mar[[eth]] <-
    map(mod_pgs_x_mar[[eth]], broom::tidy, conf.int = TRUE)
  names(tidy_pgs_x_mar[[eth]]) <- pgs
  tidy_pgs_x_mar[[eth]] <-
    list_rbind(tidy_pgs_x_mar[[eth]], names_to = "pgs") |>
    filter(grepl(':', term)) |>
    mutate(
      married = str_remove(term, ".*:married")
    )

  form_pgs_x_agefb <- paste0(pgs, "*agefb") |>
    map(\(x) reformulate(c(x, "agefb", pcs), response = "rlrs"))
  mod_pgs_x_agefb[[eth]]  <-
    map(form_pgs_x_agefb, 
               \(x) svyglm_quiet(x, r_eth))

  tidy_pgs_x_agefb[[eth]] <-
    map(mod_pgs_x_agefb[[eth]], broom::tidy, conf.int = TRUE) |> 
    setNames(pgs) |> 
    list_rbind(names_to = "pgs") |>
    filter(grepl(':', term)) |>
    mutate(
      agefb = str_remove(term, ".*:agefb")
    ) 
}

@@

$$
\hat{\beta}_{TRUE} = \hat{\beta}_{PGS}\sqrt{\frac{h^2}{R^2_{PGS}}}
$$
